Resource Allocation in an Open RAN System using Network Slicing\vspace{-.1cm}

### 2021-07-31 Latex/journal_v1/V5/pr.tex
add figure

In this section, we describe the downlink system in O-RAN slicing as depicted in Figure \ref{fig:c11}. 
Here, firstly, we present the system model. Then, we obtain achievable data rates, power of O-RU and the fronthaul capacity for the downlink (DL) of the ORAN system. Afterward, we discuss about the mean delay and the power of VNFs.
Finally, the main problem is expressed.
\subsection{System Model}
Suppose we have three service types includes mMTC, eMBB and URLLC which support different applications.

Assume we have $S_1$, $S_2$ and $S_3$ different applications for the first, second and third service type, respectively ($S = S_1 + S_2 + S_3$).
So, we have $S$ preallocated slices serving these $S$ services; There are $S_1$ slices for the first service type (eMBB), $S_2$ slices for the second service type (URLLC) and $S_3$ slices for the third service type (mMTC). So each service request $s$ served by its corresponding slice.

Each Service $s_j\in \{1,2,...,S_j\} $ consists of $U_{s}$ request from the 
single-antenna UEs which require certain QoS to be able to use the requested program($j \in \{1,2,3\}$ indicate service type).
There are different application request which fall into one of these service categories. Each application request requires specific QoS. Based on the request for the application and QoS, UE may be admitted and allocated to the resources.
Each slice $s_j \in \{1,2,...,S_j \}$, $j \in \{1,2\}$ consists of  preallocated virtual resource blocks that are mapped to the Physical Resource Blocks (PRBs), $M_s^{d}$ VNFs for the processing of O-DU, $M_s^{c}$ VNFs for the processing of O-CU-UP and $M_s^{u}$ VNFs for the processing of UPF.

All $K$ PRBs can be assigned to the all UE in each service.
Also, each VNF instance is running on the virtual machine (VM) that are using resources from the data centers. Each VM, requires enough resources of CPU, memory, storage and network bandwidth.

In addition, there are $R$ multi-antenna O-RU that are shared between slices. Each O-RU $r \in \{1,2,...,R \}$
has $J$ antenna for transmitting and receiving data.
Also $\mathcal{R} = \{ r | r\in 1,2,...,R \}$ depicts the set of O-RUs.
Moreover, all O-RUs, have access to the all PRBs.
%Coordinated multipoint (CoMP) system is considered here. 
\subsection{The Achievable Rate}
The SNR of $i^{th}$ UE served at slice $s$ on PRB $k$ is obtained from
\begin{equation}\label{eq1}
\rho_{r,u(s,i)}^{k} =  \frac{|p_{r,u(s,i)}^{k}{\bold{h}_{r,u(s,i)}^{H \: k}} \bold{w}_{r,u(s,i)}^{k}|^2}{BN_0 + I_{r,u(s,i)}^{k}},
\end{equation} 
where $p_{r,u(s,i)}^{k}$ represents the transmission power from O-RU $r$ to $i^{th}$ UE served at slice $s$ on PRB $k$. 
${\bold{h}_{r,u(s,i)}^{k}} \in \mathbb{C}^{J}$ is the vector of channel gain of a wireless link from 
$r^{th}$ O-RU to the $i^{th}$ UE in $s^{th}$ slice. In addition, $\bold{w}_{r,u(s,i)}^{k} \in \mathbb{C}^{J}$ depicts the  transmit beamforming vector from $r^{th}$ O-RU to the $i^{th}$ UE in $s^{th}$ slice that is the zero forcing beamforming vector to minimize the interference which is indicated as below
\begin{equation}
\bold{w}_{r,u(s,i)}^{k} = {\bold{h}_{r,u(s,i)}^{k}}({\bold{h}_{r,u(s,i)}^{H \: k}} {\bold{h}_{r,u(s,i)}^{k}})^{-1}
\end{equation}
Moreover, $g_{u(s,i)}^r \in \{0,1\}$ is the binary variable that illustrates whether O-RU $r$ served the $i^{th}$ UE that is allocated to $s^{th}$ slice or not. 
%Here, it is assumed that each UE is served by the nearest o-RU
Also, $BN_0$ denotes the power of Gaussian additive noise, and $I_{r,u(s,i)}^{k}$ is the power of interfering signals represented as follow

@@

The achievable data rate for the $i^{th}$ UE request in the $s_{1}^{th}$ application of service type 1 (eMBB) can be written as $\mathcal{R}_{u(s_1,i)}$ that is formulated as below.
\begin{equation}\label{eq3}
\begin{split}
\mathcal{{R}}_{r,u(s_1,i)}^{k} &=  B \log_2({1+ \rho_{r,u(s_1,i)}^{k}}) ,\\
\mathcal{R}_{u(s_1,i)}^{r} &= \sum_{k=1}^{K} B \log_2({1+ \rho_{r,u(s_1,i)}^{k}} e^k_{r,u(s_1,i)}),\\
\mathcal{R}_{u(s_1,i)} &= \sum_{r=1}^{R}\mathcal{R}_{u(s_1,i)}^{r} g^r_{u(s_1,i)}
\end{split}
\end{equation}
where $B$ is the bandwidth of system. 
$\mathcal{R}_{u(s_1,i)}^{r}$ is the achievable rate of each RU $r$ to UE $i$ in slice $s_1$.
Since the blocklength in URLLC and mMTC is finite, the achievable data rate for the $i^{th}$ UE request in the $s_{j}^{th}$ ( $j \in \{2,3\}$) application of service type 2 (URLLC) and 3 (mMTC) is not achieved from Shannon Capacity formula. So, for the short packet transmission the achievable data rate is approximated from follow
\begin{equation}\label{eq11}
\begin{split}
\mathcal{{R}}_{r,u(s_j,i)}^{k} &= B \log_2({1+ \rho_{r,u(s_j,i)}^{k}} - \zeta_{u(s_j,i)}^{k}){e}_{u(s_j,i)}^{k},\\
\mathcal{R}_{u(s_j,i)}^{r} &= \sum_{k=1}^{K} B (\log_2({1+ \rho_{u(s_2,i)}^{k}})- \zeta_{u(s_j,i)}^{k}){e}_{u(s_j,i)}^{k}\\
\mathcal{R}_{u(s_j,i)} &= \sum_{r=1}^{R}\mathcal{R}_{u(s_j,i)}^{r} g^r_{u(s_j,i)}
\end{split}
\end{equation}
Where $j \in \{1,2\}$. Also we have
\begin{equation}\label{shortPacket}
 \zeta_{u(s_j,i)}^{k} = log_2({e})Q^{-1}(\epsilon) \sqrt{\frac{\mathfrak{C}_{u(s_j,i)}^{k}}{N_{u(s_j,i)}^{k}}})
\end{equation}
Where, $\epsilon $ is the transmission probability, $Q^{-1}$ is the inverse of Q- function (Gaussian),
$\mathfrak{C}_{u(s_j,i)}^{k} = 1 - \frac{1}{(1+\rho_{u(s_j,i)}^{k})^2}$ depicts the channel dispersion of UE  $i$ at slice $s_j$, experiencing PRB $k$ and
$N_{u(s_j,i)}^{k}$ represents the blocklength of it. 
$\mathcal{R}_{u(s_j,i)}^{e,r}$ is the achievable rate of each O-RU $r$ to UE $i$ in slice $s_j$.

@@

Let $P_r$ denote the power of transmitted signal from the $r^{th}$ O-RU to UEs served by it.
From \eqref{eq2}, we have,

@@

Since we have fiber link between O-RU and O-DU, the rate of users on the fronthual link between O-DU and the $r^{th}$ O-RU  is formulated as

@@

In this part, the end to end mean delay for a service is obtained.

@@

Total delay is sum of processing delay, transmission delay and propagation delay. 
The propagation delay is the time takes for a signal to reach to its destination. So it has a constant value based on the length of fiber link ($T = L/c$, where $L$ is the length of link and c is the speed of signal).

@@

($T = \frac{\alpha}{R}$ Where, $R$ is the rate of transmission in each link and $\alpha$ is the mean arrival data rate of the each link which is constant in this model.)

@@

Assume the mean arrival data rate of the UPF layer for slice $s$ ($\alpha_{s}^U$) is approximately equal to the mean arrival data rate of the O-CU-UP layer ($\alpha_{s}^C$) and O-DU ($\alpha_{s}^D$). so $\alpha_{s} =\alpha_{s}^U \approx \alpha_{s}^C \approx \alpha_{s}^D$. since, by using Burke’s Theorem, the mean arrival data rate of the second and third layer which are processed in the first layer is still Poisson with rate $\alpha_{s}$.

@@

Each packet is processed by one of the VNFs of a slice. So, the mean delay for the $s^{th}$ slice in the first and the second layer, modeled as M/M/1 queue, is formulated as follow, respectively

@@

Moreover, $1/\mu_s^d$, $1/\mu_s^c$ and $1/\mu_s^u$ are the mean service time of the O-DU, O-CU and the UPF layers respectively.

@@

$\phi_{s}$. So the total power of the system for all active DCs that are connected to slices can be represented as

@@

Moreover, $\phi_s^u$, $\phi_s^c$ and $\phi_s^d$ are the static cost of energy in UPF, O-CU and O-DU, respectively. 

@@

The aim of this paper is to maximize the sum rate of all UEs with the presence of constraints which is written as follow,

@@

\eqref{c11}, and \eqref{c12}, indicate that the power of each RU do not exceed the maximum power, and the power of each UE is a positive integer value, respectively. 
Also \eqref{c13} shows that the rate of each UE requesting eMBB, URLLC and mMTC is more than a threshold, respectively.
\eqref{c15} and \eqref{c16} expressed the limited capacity of the fronthaul link, and the limited delay of receiving signal, respectively.

@@

%\eqref{c16} is a reliability condition that the delay in each layer should be less than threshold.
\eqref{c17} and \eqref{c18-1} guarantee that O-RU and PRB is associated to the UE, respectively.
Also, \eqref{c18} ensure that each PRB can not be assigned to more than one UE associated to the same O-RU.
In addition, \eqref{c19} indicate that the static cost of energy of VNFs in each slice do not exceed from the threshold. 

@@

In this section, we first apply some simplifications to the system; Solving problem \eqref{problem} is complicated due to the fact that this problem is 
a non-convex problem and it is a 
mixed integer non-linear problem (MINLP) with a binary variable and an integer variable. 
In the following, we apply the simplifications to reformulate MINL parts and use iterative heuristic algorithm to solve the reformulated problem.
We solve this problem in two level iteratively until it converges; In the first level,
parameters ($\boldsymbol{P}, \boldsymbol{E}, \boldsymbol{M}$) are obtained by relaxing and reformulating parameters and turn it to convex problem; Afterward we solve it by dual optimization problem.
In the second level, finding optimal O-RU association ($ \boldsymbol{G}$) is concerned with the fixed parameter of power, PRB allocation and number of VNFs.   

@@

Therefore, we just need to have constraint below

@@

Thus the user in each slice $s$ that have the largest value of $\mathfrak{F}^{k}_{r,u(s,i)}$, should be allocated to the PRB $k$; Due to the fact that just one PRB can be allocated to a UE between those UEs (regardless to the services) that are associated to the same O-RU.

@@

Assume $\boldsymbol{P}$ and $\boldsymbol{E}$ are fixed, we want to find $\boldsymbol{G}$.
Next we introduce a greedy algorithm that assign one O-RU to each UE.

@@

We solve this problem using GAAOU which is a greedy algorithm \eqref{alg1} as follow \cite{akccay2007greedy,lee2018dynamic}.

@@

Afterward, we assign the O-RU that provides the highest achievable rate for each UE (we start from the UEs on the slices with highest priority) on the condition that 
it does not exceed the value of each O-RU (that is a function of maximum power and capacity of O-RU).

@@

Firstly, we fixed $\boldsymbol{G}$, then $\boldsymbol{P}$ and $\boldsymbol{E}$ is achieved using Lagrangian method.  
Afterward, $\boldsymbol{G}$ is updated using GAAOU algorithm. This process is repeated until it converges. 
The whole algorithm is depicted as follow (Algorithm \eqref{alg2}).

>>>>

In this section, we describe the downlink system in O-RAN slicing as depicted in Figure \ref{fig:c11}. 
Here, firstly, we present the system model. Then, we obtain achievable data rates, power of O-RU, and the fronthaul capacity for the downlink (DL) of the ORAN system. Afterward, we discuss the mean delay and the power of VNFs.
Finally, the main problem is expressed.
\subsection{System Model}
Suppose we have three service types includes mMTC, eMBB, and URLLC, which support different applications.

Assume we have $S_1$, $S_2$ and $S_3$ different applications for the first, second and third service type, respectively ($S = S_1 + S_2 + S_3$).
So, we have $S$ preallocated slices serving these $S$ services; There are $S_1$ slices for the first service type (eMBB), $S_2$ slices for the second service type (URLLC), and $S_3$ slices for the third service type (mMTC). So each service request $s$ is served by its corresponding slice.

Each Service $s_j\in \{1,2,...,S_j\} $ consists of $U_{s}$ request from the single-antenna UEs which require certain QoS to be able to use the requested program($j \in \{1,2,3\}$ indicate service type).
There are different application requests which fall into one of these service categories. Each application request requires a specific QoS. Based on the application and QoS request, UE may be admitted and allocated to the resources.
Each slice $s_j \in \{1,2,...,S_j \}$, $j \in \{1,2\}$ consists of  preallocated virtual resource blocks that are mapped to the Physical Resource Blocks (PRBs), $M_s^{d}$ VNFs for the processing of O-DU, $M_s^{c}$ VNFs for the processing of O-CU-UP and $M_s^{u}$ VNFs for the processing of UPF.

All $K$ PRBs can be assigned to all UE in each service.
Also, each VNF instance runs on a virtual machine (VM) that uses resources from the data centers. Each VM requires enough resources of CPU, memory, storage, and network bandwidth.

In addition, there are $R$ multi-antenna O-RU that are shared between slices. Each O-RU $r \in \{1,2,...,R \}$
has $J$ antenna for transmitting and receiving data. Also $\mathcal{R} = \{ r | r\in 1,2,...,R \}$ depicts the set of O-RUs. Moreover, all O-RUs, have access to all PRBs.

%Coordinated multipoint (CoMP) system is considered here. 
\subsection{The Achievable Rate}
The SNR of $i^{th}$ UE served at slice $s$ on PRB $k$ is obtained from
\begin{equation}\label{eq1}
\rho_{r,u(s,i)}^{k} =  \frac{|p_{r,u(s,i)}^{k}{\bold{h}_{r,u(s,i)}^{H \: k}} \bold{w}_{r,u(s,i)}^{k}|^2}{BN_0 + I_{r,u(s,i)}^{k}},
\end{equation} 
where $p_{r,u(s,i)}^{k}$ represents the transmission power from O-RU $r$ to $i^{th}$ UE served at slice $s$ on PRB $k$. 
${\bold{h}_{r,u(s,i)}^{k}} \in \mathbb{C}^{J}$ is the vector of channel gain of a wireless link from 
$r^{th}$ O-RU to the $i^{th}$ UE in $s^{th}$ slice. In addition, $\bold{w}_{r,u(s,i)}^{k} \in \mathbb{C}^{J}$ depicts the  transmit beamforming vector from $r^{th}$ O-RU to the $i^{th}$ UE in $s^{th}$ slice that is the zero forcing beamforming vector to minimize the interference which is indicated as below
\begin{equation}
\bold{w}_{r,u(s,i)}^{k} = {\bold{h}_{r,u(s,i)}^{k}}({\bold{h}_{r,u(s,i)}^{H \: k}} {\bold{h}_{r,u(s,i)}^{k}})^{-1}
\end{equation}
Moreover, $g_{u(s,i)}^r \in \{0,1\}$ is the binary variable that illustrates whether O-RU $r$ served the $i^{th}$ UE that is allocated to $s^{th}$ slice or not. 
%Here, it is assumed that each UE is served by the nearest o-RU
Also, $BN_0$ denotes the power of Gaussian additive noise, and $I_{r,u(s,i)}^{k}$ is the power of interfering signals represented as follow.

@@

The achievable data rate for the $i^{th}$ UE request in the $s_{1}^{th}$ application of service type 1 (eMBB) can be written as $\mathcal{R}_{u(s_1,i)}$ that is formulated as below.
\begin{equation}\label{eq3}
\begin{split}
\mathcal{{R}}_{r,u(s_1,i)}^{k} &=  B \log_2({1+ \rho_{r,u(s_1,i)}^{k}}) ,\\
\mathcal{R}_{u(s_1,i)}^{r} &= \sum_{k=1}^{K} B \log_2({1+ \rho_{r,u(s_1,i)}^{k}} e^k_{r,u(s_1,i)}),\\
\mathcal{R}_{u(s_1,i)} &= \sum_{r=1}^{R}\mathcal{R}_{u(s_1,i)}^{r} g^r_{u(s_1,i)}
\end{split}
\end{equation}
where $B$ is the bandwidth of system. 
$\mathcal{R}_{u(s_1,i)}^{r}$ is the achievable rate of each RU $r$ to UE $i$ in slice $s_1$.
Since the blocklength in URLLC and mMTC is finite, the achievable data rate for the $i^{th}$ UE request in the $s_{j}^{th}$ ( $j \in \{2,3\}$) application of service type 2 (URLLC) and 3 (mMTC) is not achieved from Shannon Capacity formula. So, for the short packet transmission, the achievable data rate is approximated from following
\begin{equation}\label{eq11}
\begin{split}
\mathcal{{R}}_{r,u(s_j,i)}^{k} &= B \log_2({1+ \rho_{r,u(s_j,i)}^{k}} - \zeta_{u(s_j,i)}^{k}){e}_{u(s_j,i)}^{k},\\
\mathcal{R}_{u(s_j,i)}^{r} &= \sum_{k=1}^{K} B (\log_2({1+ \rho_{u(s_2,i)}^{k}})- \zeta_{u(s_j,i)}^{k}){e}_{u(s_j,i)}^{k}\\
\mathcal{R}_{u(s_j,i)} &= \sum_{r=1}^{R}\mathcal{R}_{u(s_j,i)}^{r} g^r_{u(s_j,i)}
\end{split}
\end{equation}
Where $j \in \{1,2\}$. Also we have
\begin{equation}\label{shortPacket}
 \zeta_{u(s_j,i)}^{k} = log_2({e})Q^{-1}(\epsilon) \sqrt{\frac{\mathfrak{C}_{u(s_j,i)}^{k}}{N_{u(s_j,i)}^{k}}})
\end{equation}
Where, $\epsilon $ is the transmission probability, $Q^{-1}$ is the inverse of Q- function (Gaussian),
$\mathfrak{C}_{u(s_j,i)}^{k} = 1 - \frac{1}{(1+\rho_{u(s_j,i)}^{k})^2}$ depicts the channel dispersion of UE  $i$ at slice $s_j$, experiencing PRB $k$ and
$N_{u(s_j,i)}^{k}$ represents the blocklength of it. 
$\mathcal{R}_{u(s_j,i)}^{e,r}$ is the achievable rate of each O-RU $r$ to UE $i$ in slice $s_j$.

@@

Let $P_r$ denote the power of the transmitted signal from the $r^{th}$ O-RU to UEs served by it. From \eqref{eq2}, we have,

@@

Since we have a fiber link between O-RU and O-DU, the rate of users on the fronthual link between O-DU and the $r^{th}$ O-RU  is formulated as

@@

In this part, the end-to-end mean delay for a service is obtained.

@@

Total delay is the sum of processing delay, transmission delay, and propagation delay. 
The propagation delay is the time takes for a signal to reach its destination. So it has a constant value based on the length of the fiber link ($T = L/c$, where $L$ is the length of the link and c is the speed of signal).

@@

($T = \frac{\alpha}{R}$ Where $R$ is the rate of transmission in each link and $\alpha$ is the mean arrival data rate of each link which is constant in this model.)

@@

Assume the mean arrival data rate of the UPF layer for slice $s$ ($\alpha_{s}^U$) is approximately equal to the mean arrival data rate of the O-CU-UP layer ($\alpha_{s}^C$) and O-DU ($\alpha_{s}^D$). so $\alpha_{s} =\alpha_{s}^U \approx \alpha_{s}^C \approx \alpha_{s}^D$. since, by using Burke’s Theorem, the mean arrival data rate of the second and third layers, which are processed in the first layer, is still Poisson with rate $\alpha_{s}$

@@

Each packet is processed by one of the VNFs of a slice. So, the mean delay for the $s^{th}$ slice in the first and the second layer, modeled as M/M/1 queue, is formulated as follows, respectively.

@@

Moreover, $1/\mu_s^d$, $1/\mu_s^c$, and $1/\mu_s^u$ are the mean service time of the O-DU, O-CU, and the UPF layers, respectively.

@@

$\phi_{s}$. So the system's total power for all active DCs connected to slices can be represented as follows.

@@

Moreover, $\phi_s^u$, $\phi_s^c$, and $\phi_s^d$ are the fixed cost of energy in UPF, O-CU, and O-DU, respectively. 

@@

This paper aims to maximize the sum rate of all UEs with the presence of constraints as follows.

@@

\eqref{c11}, and \eqref{c12}, indicate that the power of each RU does not exceed the maximum power, and the power of each UE is a positive integer value, respectively.  
Also, \eqref{c13} shows that the rate of each UE requesting eMBB, URLLC, and mMTC is more than a threshold, respectively.
\eqref{c15} and \eqref{c16} expressed the limited capacity of the fronthaul link and the limited delay of receiving the signal, respectively.

@@

%\eqref{c16} is a reliability condition that the delay in each layer should be less than the threshold.
\eqref{c17} and \eqref{c18-1} guarantee that O-RU and PRB are associated with the UE, respectively.
Also, \eqref{c18} ensures that each PRB can not be assigned to more than one UE associated with the same O-RU.
In addition, \eqref{c19} indicates that the fixed cost of energy of VNFs in each slice does not exceed the threshold. 

@@

In this section, we first apply some simplifications to the system; Solving problem \eqref{problem} is complicated since this problem is non-convex and it is a  mixed-integer non-linear problem (MINLP) with a binary variable and an integer variable. 
We simplified and reformulated MINL parts and used an iterative heuristic algorithm to solve the problem.
We solve this problem in two-level iteratively until it converges. In the first level, parameters ($\boldsymbol{P}, \boldsymbol{E}, \boldsymbol{M}$) are obtained by relaxing and reformulating parameters and turn them into a convex problem; Afterward, we solve it by dual optimization problem.
In the second level, finding optimal O-RU association ($ \boldsymbol{G}$) is concerned with the fixed parameter of power, PRB allocation, and number of VNFs.   

@@

Therefore, we need to have constraints below

@@

Thus the user in each slice $s$ that have the the most considerable value of $\mathfrak{F}^{k}_{r,u(s,i)}$, should be allocated to the PRB $k$. Since just one PRB can be allocated to a UE between those UEs (regardless to the services) that are associated to the same O-RU.

@@

Assume $\boldsymbol{P}$ and $\boldsymbol{E}$ are fixed, we want to find $\boldsymbol{G}$.
Next, we introduce a greedy algorithm that assigns one O-RU to each UE.

@@

We solve this problem using GAAOU, which is a greedy algorithm \eqref{alg1} as follows \cite{akccay2007greedy,lee2018dynamic}.

@@

Afterward, we assign the O-RU that provides the highest achievable rate for each UE (we start from the UEs on the slices with the highest priority) on the condition that 
it does not exceed the value of each O-RU (that is, a function of maximum power and capacity of O-RU).

@@

Firstly, we fixed $\boldsymbol{G}$, then $\boldsymbol{P}$ and $\boldsymbol{E}$ is achieved using the Lagrangian method.  
Afterward, $\boldsymbol{G}$ is updated using the GAAOU algorithm. This process is repeated until it converges. 
The whole algorithm is depicted as follows (Algorithm \eqref{alg2}).
